### 2014-11-12 docs/ms.Rmd
Tightening the intro - in progress

# The transparency of methods-reporting in studies measuring biological colouration

@@

## Introduction
Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration. For these biologists, identifying best practices in measuring color and color pattern and in determining how other species see colour has been a great challenge. Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography and visual modeling.

The application of spectrophotometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966]. Before the widespread use of spectrophotometry, researchers assigned categorical color descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although when applied carefully they maybe useful, these techniques in general have limited progress in the field. For example, many researchers designed color assessment methods specifically for their projects [e.g. @aldrich_ecogeographic_1991] rendering data by different researchers incomparable. This was/is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX). Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994](MORE). Human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

@@

Sensory models typically attempt to describe the subjective perception of color or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g. @endler_comparing_2005; @vorobyev_receptor_1998]. Furthermore, an animal’s ability to distinguish between reflectance spectra can then be predicted based on the differential responses of a visual system to varying reflectance spectra. This approach enables researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.  
Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus log-transformed), the use of any color constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of model species, such as honeybees, and so assumptions and extrapolations are necessary when models are applied to related species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].

>>>>

# The transparency of methods-reporting in studies of biological colouration

@@

## Introduction
Generations of biologists have endeavored to explain the mechanisms, functions, and evolution of animal and plant coloration [@poulton_colours_1890; @thayer_concealing_1909; @wallace_natural_1891]. Identifying best practices for measuring and analysing colors and patterns has been an ongoing challenge. Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (variously known as spectroradiometry, spectroscopy, ?), photography and, more recently, visual modeling (CITES).

The application of spectrophotometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966]. Before the widespread use of spectrophotometry, researchers assigned categorical color descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on subjective human assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although potentially useful when used carefully, the persistence of these techniques in general has limited progress in the field. For example, many researchers designed color assessment methods specifically for their projects [e.g. @aldrich_ecogeographic_1991] rendering data by different researchers incomparable. This is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX). Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994](MORE). Human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

@@

Sensory models typically attempt to describe the subjective perception of color or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g. @endler_comparing_2005; @vorobyev_receptor_1998]. Furthermore, an animal’s ability to distinguish between reflectance spectra can then be predicted based on the differential responses of a visual system to varying reflectance spectra. This approach enables researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.

Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus log-transformed), the use of any color constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of model species, such as honeybees, and so assumptions and extrapolations are necessary when models are applied to related species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].
